### 2022-04-13 greta/mcmc.Rmd
10 replicates

---
output: github_document
---

```{r}
library(tidyverse)
set.seed(1234567)

np.clip <- function(x, a, b) {
  if(x < a) return(a)
  if(x > b) return(b)
  x
}

# View the bistable curve: 
# curve( (0.75 - x)**2 * (0.25 - x), 0, 1)

step <- function(N, eta,  a = 0.75, b = 0.25) {
  N <- N + (a - N)**2 * (b - N) + eta
  N <- np.clip(N, 0, 1)
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     mu =0, # -5e-4,
                     sigma = 1e-2,
                     a = 0.75, 
                     b = 0.25) {
  eta <- rnorm(t_max, mu, sigma)
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], a = a, b = b)
  }
  tibble::tibble(t = 1:t_max, N = N)
}

sims <- purrr::map_dfr(1:10, \(i) simulate(), .id = "i")
sims |> ggplot(aes(t, N, group=i)) + geom_line(alpha=0.06)
```

```{r}
gsims <- sims |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))

x_t <- gsims$N
x_t1 <- gsims$xt1

library(greta)
a <- uniform(0, 10)
b <- uniform(0, 10)
sigma <- uniform(0, 10)
mean <-  x_t + (a - x_t) ^ 2 * (b - x_t)
distribution(x_t1) <- normal(mean, sigma )
m <- model(a, b, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("mcmc_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 10000, warmup = 5000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(100)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate() ,.id = "i")

bind_rows(sims, posterior_sims, .id = "model")|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.3)

```

# Scoring

First, a single 'observed' sample from the true:

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- posterior_sims |> pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> select(-t) |> as.matrix()

# Score just the 1st replicate "true" observations
observed <- sims |> filter(i == 1) |> pull(N)
scores(observed, dat)

```
```{r}
# score over all replicates:

rep_scores <- sims |> group_by(i) |> group_map(~ scores(.x$N, dat)) |> bind_rows()
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps)) + 
(ggplot(rep_scores) + geom_histogram(aes(logs)) )

```