### 2022-04-19 greta/hopf_mcmc.Rmd

---
output: github_document
---

```{r}
library(tidyverse)
library(greta)
library(bayesplot)

set.seed(4242)
train_reps <- 10
train_t_max <- 100
test_t_max <- 100
test_reps <- 100

np.clip <- function(x, a, b) {
  if(x < a) return(a)
  if(x > b) return(b)
  x
}
```

Cycles start around `K >= 21`, though some stochastic resonance visible before then.
`K` will go linearly from 15 to 25 with `delta = 0.1`.
Make delta negative and start with a bigger `Ko` to run in reverse.

```{r}
step <- function(H, P, eta,  t=0, 
                 p = list(r = 0.75, c = 0.25, Ko = 15, delta=0.1)) {
  K <- p$Ko + p$delta*t
  N <- numeric(2)
  N[1] <- H * exp(p$r * (1 - H/K) - p$c * P + eta[1])
  N[2] <- H * exp(p$r * (1 - H/K) ) * (1 - exp(-p$c * P+ eta[2]) )
  N[1] <- np.clip(N[1], 0, 100)
  N[2] <- np.clip(N[2], 0, 100)
  N
}

simulate <- function(t_max = 100, 
                     p = list(H_init = 9, P_init = 1,
                              r = 0.75, c = 0.1,  Ko = 15, delta = 0.1,
                              sigma_H = 2e-2, sigma_P = 2e-2)
                     ) {
  eta <- array(c(rnorm(t_max, 0, p$sigma_H), 
                 rnorm(t_max, 0, p$sigma_P)),
               dim=c(t_max,2))
  N   <- array(NA, c(t_max,2))

  N[1,] <- c(p$H_init, p$P_init)
  for (t in 1:(t_max-1)) {
    N[t+1,] <- step(N[t,1], N[t,2], eta[t,], t=t, p = p)
  }
  tibble::tibble(t = 1:t_max, H = N[,1], P = N[,2])
}
```

```{r}
t_max <- train_t_max + test_t_max
 p = list(H_init = 9, P_init = 1,
          r = 0.75, c = 0.1,  Ko = 14, delta = 0.08,
          sigma_H = 2e-2, sigma_P = 2e-2)
sim <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=t_max, p = p), .id = "i")
train <- sim |> filter(t <= train_t_max)
test <- sim |> filter(t > train_t_max)
# test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

## show training data
train |> ggplot(aes(H, P, group=i)) + geom_path(alpha=0.4)

sim |> 
  pivot_longer(c(P,H), values_to="density", names_to="species") |>
  ggplot(aes(t, density, col=species, group=interaction(i,species))) + geom_line(alpha=0.2) +
  geom_vline(aes(xintercept = train_t_max))
```

```{r}
gsims <- train |> 
  group_by(i) |> 
  mutate(H1 = lead(H),
         P1 = lead(P)) |> 
  filter(t<max(t)) |> 
  ungroup()

stopifnot(all(gsims$P1 > 0))

H_t  <- gsims$H 
H_t1 <- log( gsims$H1 ) 
P_t  <- gsims$P
P_t1 <- log( gsims$P1 )
t <- gsims$t
```

```{r}
library(greta)
r <- uniform(0, 10)
c <- uniform(0, 10)
Ko <- uniform(0, 50)
delta <- uniform(-1,1)
sigma_H <- uniform(0, 1)
sigma_P <- uniform(0, 1)

K <- Ko + delta * t
mu_H <- log( H_t ) +  (r * (1 - H_t/K) - c * P_t) 
distribution(H_t1) <- normal(mu_H, sigma_H )

mu_P <- log( H_t ) +  (r * (1 - H_t/K) ) + log(1 - exp(-c * P_t) )
distribution(P_t1) <- normal(mu_P, sigma_P )
m <- model(r, c, Ko, delta, sigma_H, sigma_P)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("hopf_cache"))
bench::bench_time({                 
    draws <- mmcmc(m, 
                 n_samples = 10000, warmup = 5000,
                 chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
#test_reps <- 100
## draw test_reps number of samples
left_off <- train |> 
  filter(t == train_t_max) |> 
  select(H,P) |> 
  rename(H_init = H, P_init = P) |>
  slice_sample(n = test_reps, replace=TRUE)

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps) |>
  bind_cols(left_off)

posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max,
                               p = q
                               ) ,.id = "i") |>
  mutate(t = t+train_t_max)
```

```{r}
bind_rows(mutate(train, type="training"), 
          mutate(test,type="observed"),  
          mutate(posterior_sims, type="predicted")) |>
  rename("host" = H, "parasite" = P) |>
pivot_longer(c(host, parasite), values_to="density", names_to="species") |>
  ggplot(aes(t, density, col=type, group=interaction(i,type))) + geom_line(alpha=0.2) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```

```{r}
true <- as_tibble(p) |> select(-H_init, -P_init)  |> gather(variable, value)

bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

# Scoring

First, a single 'observed' sample from the true:

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "P") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$P, dat)) |> 
  bind_rows()
})
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps)) + 
(ggplot(rep_scores) + geom_histogram(aes(logs)) )

```