### 2022-04-22 greta/sn_mcmc.Rmd
sn mcmc :turtle:

---
output: github_document
---

```{r}
library(tidyverse)
library(greta)
library(bayesplot)

set.seed(4242)
train_reps <- 1
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

np.clip <- function(x, a, b) {
  if(x < a) return(a)
  if(x > b) return(b)
  x
}
```

```{r}
step <- function(N, 
                 eta,
                 t,
                 pars) {

        h <- np.clip(pars$h_init + pars$alpha * t, 0, 0.27)
        N <- N + pars$r * N * (1 - N / pars$K) - 
          h * (N**2 / (pars$s**2 + N**2)) + eta
        N <- np.clip(N, 0, 1)
        N
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     pars = list(
                        r = 1,
                        K = 1,
                        s = 0.1,
                        h_init = .15,
                        alpha = 0.0015,
                        mu = 0,
                        sigma = 0.00
                      )
                    ) {
  eta <- rnorm(t_max, pars$mu, pars$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], t, pars)
  }
  tibble::tibble(t = 1:t_max, N = N)
}
```

```{r}
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02
          )
t_max <- train_t_max + test_t_max
sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")

train <- sims |> filter(t <= train_t_max)
test <- sims |> filter(t > train_t_max)
```

```{r}
sims <- bind_rows(test, train, .id = "data")
sims |> 
  ggplot(aes(t, N, group=interaction(i,data), col=data)) + 
  geom_line()
```

```{r}
gsims <- train |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))
x_t <- gsims$N
x_t1 <- gsims$xt1
t <- gsims$t

library(greta)
r <- uniform(0, 10)
K <- uniform(0, 10)
s <- uniform(0,10)
h_init <- uniform(0,10)
alpha <- uniform(0, 10)
sigma <- uniform(0, 10)
h <- h_init + alpha * t
mean <-  x_t + r*x_t*(1 - x_t / K) - h*(x_t^2 / (s^2 + x_t^2))
distribution(x_t1) <- normal(mean, sigma)
m <- model(r, K, s, h_init, alpha, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("sn_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
## draw test_reps number of samples

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max, p = q) ,.id = "i")

bind_rows(
  mutate(test, model="true"), 
  mutate(posterior_sims, model="predicted")
)|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.1)

```

```{r}
true <- tibble(a = 0.75, b = 0.25, sigma = 1e-2) |> gather(variable, value)
bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

# Scoring

First, a single 'observed' sample from the true:

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()
})
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps), bins=20) + 
(ggplot(rep_scores) + geom_histogram(aes(logs), bins=20) )

```