### 2022-04-23 greta/mcmc.Rmd
generalizing functions a bit

---
output: github_document
---

```{r}
library(tidyverse)
library(greta)
library(bayesplot)
source("R/saddle_node.R")
```

```{r}
set.seed(4242)
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02,
            t_init = 0
          )
t_max <- train_t_max + test_t_max

simulate <- simulate_sn
```

```{r}
train_reps <- 100
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")
train <- sims |> filter(t <= train_t_max)
test <- sims |> filter(t > train_t_max)
```

```{r}
m <- greta_model_sn(train)

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 100000, warmup = 80000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
plot_posteriors(draws, pars)
```

```{r}
fc <- forecast(draws, train, test, simulate)

fc |> ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.2)

```

# Scoring

First, a single 'observed' sample from the true:

```{r}

# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()
})
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps), bins=20) + 
(ggplot(rep_scores) + geom_histogram(aes(logs), bins=20) )

```