### 2022-05-13 manuscript/manuscript.Rmd

---
title: "Ecological Limits to Forecasting: Estimating Uncertainty in Critical Transitions with Deep Learning"
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. Meaningful representations of uncertainty are an essential component of effective ecological forecasting.
  2. Machine learning
  3. We examine scenarios in which stochasticity and non-linearity can lead to extreme uncertainty under common ecological models.  
  4. We show that ..

keywords:
  - Time Series
  - Forecasting
  - Machine Learning
  - Artificial Intelligence
  - Tipping points
bibliography: references.bib
nocite: |

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
#  - \usepackage{lineno}
#  - \usepackage{endfloat}
#  - \linenumbers

output: 
  rticles::arxiv_article:
    keep_tex: true
  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(patchwork)
library(arrow)

bound <- function(x, percentile = 0.975) {
  dplyr::nth(x, percentile * length(x), order_by = x)
}
```

# Introduction

Forecasting plays an important and rapidly growing role in both testing our fundamental understanding of ecological processes, and informing ecological applications and conservation decision-making [@dietze_iterative_2018; @schindler_portfolio_2015].
Meanwhile, recent advances in machine learning have rapidly improved the prevalence and accuracy of short term forecasts in many fields [@kao_exploring_2020; @lyu_lstm_2020; @du_multivariate_2020]. 
Will these emerging methods improve the capacity for forecasts in ecological systems as well?
Ecological dynamics are notoriously complex, with uncertainty and non-linearity playing critical roles [@Boettiger2018; @hallett_why_2004; @Oveskienen2010].
These challenges are nowhere more evident then in _critical transitions_, sudden shifts in the states or patterns of ecosystem dynamics that are more important and more difficult to predict than gradual changes.
Here, we examine several of the best-known examples of critical transitions in ecological systems.
We evaluate the ability of some of the most promising machine learning methods for probabilistic forecasts relative to traditional statistical and mechanistic approaches applied to several classic models in ecology.

# Materials and Methods

<!-- some more verbose bits about bifurcation models probably belongs in discussion instead -->
We will focus the analysis on several different forecasting scenarios based around two classic models in population ecology: 
Robert May's consumer-resource model [@May1977], and the Nicholson-Bailey parasitoid-host model [@nicholson_balance_1935]. 
Though these models may appear simple when measured against high-dimensional and parameter rich models found in some management contexts such as fisheries, they can exhibit rich nonlinear dynamics and provide greater capacity to generalize [@Levins1966; @Getz2018]. 
These textbook models have been well studied and form the basis of half a century of research in ecology, including much recent work on topics such as resilience and tipping points which has had important theoretical and practical management outcomes [@Folke2004; @Fischer2009; @Polasky2011].
May's model exhibits alternative stable states. 
In this one-dimensional model, transitions between these states can occur due to intrinsic stochasticity, external forcing, or the gradual environmental change that results in a catastrophic saddle-node bifurcation and generates hysteresis. 
The Nicholson-Bailey model is a two species model which contains a Hopf bifurcation, a non-catastrophic bifurcation which either creates or destroys a limit cycle -- a stable oscillatory pattern.

Assessing the accuracy of forecasting methods in the face of such bifurcation dynamics is a particularly important question for ecological systems and global environmental change problems.
Bifurcations represent the kind of non-linear responses complex systems can make as the result of slowly changing parameters (i.e. the 'environment' in which those dynamics take place.)
This can create a particularly challenging forecasting task when such transitions have not been previously observed in the same system, requiring the forecast to anticipate dynamics for which there are no corresponding analog in the historical data to date. 
Forecast skill under such no-analog conditions may be particularly relevant to ecological forecasting in the context of global change [@Williams2007].

We provide fully reproducible coded examples in R and Python for performing, scoring, and visualizing each of the forecasts considered here.
After significant time spent considering alternative frameworks, we have emphasized those which best met our requirements for performance, ease-of-use, flexibility, and support for the latest probabilistic machine learning models for forecasting. 
Most of our forecasts use the `darts` framework, a sophisticated and well documented Python library with support for a wide range of methods.
Our model-based MCMC forecasts use the `greta` framework, an R library that uses Python-based `tensorflow` probability to achieve better performance.
While Python-based frameworks currently have the edge in performance and access modern ML algorithms, they lag behind in attention to statistical issues such as the computation of strictly proper skill scores (discussed below.)

Thus our examples of scoring and visualization will rely on a collection of R packages, in particular, `scoringRules` for the efficient calculation of Continuous Ranked Probability Score (CRPS) and logarithmic probability (Logs) scores for forecast ensembles [@gneiting_strictly_2007].
Following popular conventions, we express both skill scores in error-orientation, that is, larger values indicate worse skill (higher degree of error).

We expect greater convergence between methods available in R and Python in the future, as already illustrated in the example of `greta`.
Complete code for all examples presented here can be found at https://github.com/boettiger-lab/mee_tipping_point_forecasting.

```{r fig.cap="Forecast scenarios. A. a Hopf bifurcation: a stable node develops into a limit cycle. B. The saddle-node bifurcation."}
hopf <- read_csv("../greta/forecasts/simulation=hopf/hopf_increasing.csv.gz")
panelA <- hopf |> filter(type %in% c("historical", "true"), i < 10) |>
  rename(species = variable) |> 
  mutate(species = forcats::fct_recode(species, host = "H", parasitoid = "P")) |>
  ggplot(aes(t, value, group=interaction(species, i, type), col=species)) + 
  geom_line(show.legend = FALSE, alpha = 0.7) + theme_bw()+  scale_color_grey() +
  ggtitle("A. Hopf bifurcation")

sn <- read_csv("../greta/forecasts/simulation=saddle/saddlenode.csv.gz")
panelB <- sn |> filter(type == "historical", i == 1) |>
  bind_rows(filter(sn, type=="true")|>mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5, show.legend = FALSE) +
  scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("B. Saddle-node bifurcation")

stoch <- read_csv("../greta/forecasts/simulation=stochastic/stochastic.csv.gz")
panelC <- stoch |> filter(type == "historical", i == 1) |>
  bind_rows(filter(stoch, type=="true", i < 50) |> mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5) + scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("C. Stochastic transition")

panelA / panelB / panelC

```

## Scenario 1: Hopf Bifurcation

The Nicholson-Bailey model describes a predator-prey dynamic for the relationship of a host species and an obligate parsatiod, originally used to model the population dynamics of blowflies (_Lucilia cuprina_) [@nicholson_balance_1935; @nicholson_outline_1954; @nicholson_compensatory_1954].
We consider the form which includes density dependence in the host species, and allow for environmental stochasticity,

\begin{align}
H_{t+1} &= H_t \exp \left( r \left(1 - \tfrac{H}{K_t}\right) - c P_t + \eta_{H,t} \right) \\
P_{t+1} &= H_t \exp \left( r \left(1  \tfrac{H}{K_t}\right) \right)  \left(1 -  \exp \left(- c P_t + \eta_{P,t}\right) \right) \\
K_{t+1} &= K_t + \delta
\label{hopf}
\end{align}

Following @dakos_methods_2012, we further allow the carrying capacity of the host, $K$ to slowly increase at linear rate, which drives a Hopf bifurcation as $K$ becomes sufficiently large. 
In a Hopf bifurcation, a stable node starts an oscillatory pattern which continues to grow in amplitude as the bifurcation parameter continues to increase.
This example illustrates one of the many kinds of challenges which nonlinear phenomena pose to forecasting: the "historical" data prior to the bifurcation never exhibit the cyclical dynamics of growing amplitude that will emerge after the bifurcation occurs.
If we had used a purely deterministic model, the dynamics would be constrained to a single stable point, corresponding to a slowly changing steady-state population size of host and parasitoid populations.
However, stochasticity in this case acts as a source of some additional information about the dynamics, as the noise excites quasi-cycles which are visible in the irregular oscillations that appear significantly prior to the emergence of true limit cycles which follow the bifurcation [@boettiger_noise_2018].
Examples use the following parameters: $H_0 = 9$, $P_0 = 1$, $r = 0.75$, $c = 0.1$, $K_0 = 14$, $\delta = 0.08$, $\sigma_H = 0.02$, $\sigma_P = 0.02$.

## Scenario 2: The saddle-node bifurcation

A yet more difficult forecasting scenario is created by the saddle-node bifurcation.
May's consumer-resource model is an one-dimensional model describing the growth of a 'resource' population (e.g. herbivore) which is grazed by a consumer [@may_population_1979].
As in the Nicholson-Bailey model, in the absence of that predation, the resource population density grows under a density-dependent pattern described by a logistic function.
The resource population is also grazed by a consumer at a rate given by a Holling type III s-curve (typically used to model handling time). 
For a certain range of parameter choices, this model supports alternative stable state dynamics, and this model has been identified and employed in explaining alternative stable state dynamics in a broad range of ecological and socio-ecological systems [@scheffer_catastrophic_2001].

\begin{align}
N_{t+1} &= N_t + r N_t \left(1 - \frac{N_t}{K} \right) - \frac{h_t N_t^2}{s^2 + N_t^2} + \eta_t \\
h_{t+1} &= h_t + \alpha \\
\eta_t &\sim \mathcal{N}(0, \sigma)
\end{align}

If the environment slowly alters one of the parameters (say, the encounter efficiency, `h_t`, in our formulation), one of the stable nodes moves closer and closer to the unstable saddle point, leading to a bifurcation that destroys the stable state, leaving the system to suddenly transition to the alternative stable state.
Saddle-node bifurcations (also known as fold bifurcations) also create a phenomenon known as hysteresis, where it is not sufficient to restore the environment to the previous parameter values to recover the previous state. 
Unlike the Hopf bifurcation which exhibits a continuous transition from a stable node to a small limit cycle that then grows, the saddle-node transition is a discontinuous or so-called 'catastrophic' bifurcation.
Due both to this sudden, catastrophic nature of the transition and the difficulty in reversing the shift after it has occurred, saddle-node bifurcations have been the subject of intense study.

Tipping point dynamics have long been identified as an important but difficult challenge for forecasting [e.g. @Scheffer2001; @Folke2004]. 
Much effort in the ecological literature so far has focused on identifying any 'early warning signs' that a catastrophic bifurcation might occur at all [@Scheffer2009; @Scheffer2012] rather than more ambitious attempts to provide quantitative probabilistic forecasts of the likely distribution of waiting times before such a transition occurs.
Tipping points resulting from Saddle-node bifurcations have been demonstrated in examples ranging from laboratory microcosms [@Dai2012; @Dai2015] to whole-ecosystem experiments [@carpenter_early_2011], and postulated as a model for global change [@Barnosky2012]. 
Examples use the following parameters: $r = 1$, $K = 1$, $s = 0.1$, $h_0 = 0.15$, $\alpha = 0.000375$, $\sigma = 0.02$, $N_0 = 0.75$.

## Scenario 3: The stochastic transition

Perhaps the most difficult of all events to predict are those in which large transitions are predominately driven by a random component. 
An example of such a transition event is possible to observe in May's consumer-resource model, in which a stochastic term occasionally results in a transition between alternative stable states.
In such cases, no forecast can precisely predict when a transition will occur, but it is nonetheless possible to deduce the correct distribution of waiting times knowing the correct model. 
In the case of small noise, transitions are Poisson distributed, such that the distribution of waiting times is roughly exponential [e.g. @kampen_stochastic_1992], though post-hoc the trajectories of such transitions can be mistaken for saddle-node transitions [@boettiger_early_2012].
To consider such cases, we will again use May's alternative stable state model, though this time leaving all parameters fixed.

In this context, predicting the probability of a transition in the future based solely on observations prior to a transition occurring is essentially impossible without additional information constraining the model estimate, as such data is equally consistent with infinitely many models or parameter choices which share the same local linearization about the stable point.  
Unlike the saddle-node bifurcation, there is no slowly warping potential basin which can be detected to inform estimates. 
Thus, in this scenario, rather than considering the problem of predicting the future evolution of a single time series based only on its historical values, we consider an alternative framing of the task:
we imagine our forecaster has access to historical data from one or more comparable systems which includes a previous stochastic transition event.
Based on this data, our forecaster seeks to identify the distribution of expected transition times for analagous systems starting from the same initial condition. 
This parallels actual practice in which researchers would draw on previous examples of stochastic transitions in a system - lake-ecosystem shifts, disease emergence, changing fire regimes, [@Scheffer2001; @Folke2004].
(Note that such stochastic transitions between alternative stable states can also create oscillatory-like dynamics when stochasticity is sufficiently high enough to drive repeated transitions from one attractor to the other and back again.  In such cases, it might be reasonable to estimate a strictly forward-looking forecast of a single system, predicting the distribution of these transitions.)
Model definition is the same as May's model for the saddle node with fixed parameter $h$, values: $r = 1$, $K = 1$, $s = 0.1$, $h_0 = h = 0.26$, $\alpha = 0$, $\sigma = 0.02$, $N_0 = 0.55$.

## Method group 1: Markov Chain Monte Carlo

As a reference case, we consider the forecast produced by an MCMC estimate of model parameters, _given the true model_.
This represents an idealized case where the nature of the underlying process is known precisely.
Uncertainty comes from parameter estimates and intrinsic stochasticity specified in the model, but does not reflect any uncertainty in our knowledge of the model structure. 
Alternative model structures, even when capable of producing the same nonlinear phenomena (i.e. the same bifurcations) will give very different forecasts. 
Even alternative prior distributions of the parameters will generally yield alternate forecasts, as likelihood ridges are common to nonlinear models.
Thus, this case represents a theoretical upper bound for the performance of forecasts by techniques which do not make such strong assumptions about the underlying processes.

## Method group 2: Statistical models (ARIMA)

<!-- Key distinction here from MCMC above is not on how model parameters are estimated, but rather on approaches that aren't making strong mechanistic assumptions about process, but are merely phenomenological type models
-->

We present forecasts produced by ARIMA models as the model-free analogs to the forecasts made using parameter estimation with MCMC.
Since ARIMA models make the assumption that the future will resemble the past via ARIMA's auto-regressive and moving average components [@hyndman_forecasting_2018], these models are not well-suited for problems with complex bifurcation dynamics.
Thus, ARIMA-based forecasts should be treated as a lower bound for the performance of non-mechanistic models.
In contrast to inference with MCMC, uncertainty with ARIMA models is estimated directly from the learned parameters [@hyndman_forecasting_2018].
Since ARIMA is a commonly encountered method, we will refer readers to @hyndman_forecasting_2018 for further discussion.

## Method group 3: Machine Learning models

Over the past decade, deep learning has become very popular for a broad range of challenging time series prediction problems [@makridakis_statistical_2018].
Deep learning models are often used to make point forecasts, yet for their application to ecological time series, it will often be necessary to use multi-step, probablisitic forecasts. 
For all the deep learning models in this study, we use the same method for uncertainty estimation:
each model is trained to learn the parameters of a Poisson distribution to maximize the log likelihood. 
To compute uncertainty intervals, we draw samples from the learned distribution at each time step and report the 95% confidence interval.

A major nuisance with deep learning methods is their instability to hyperparameters and initialization seeds [@madhyastha_model_2019].
We found that for the same set of hyperparameters, we could produce starkly different forecasts if we trained two models with different initialization seeds [@figure_].
One explanation for this instability is that machine learning models often get stuck on the local optima of loss surfaces [@madhyastha_model_2019]. 
Another likely cause is that machine learning models commonly overfit the training data [@mehta_high-bias_2019].
Across deep learning, overfitting is a fundamental issue, arising from neural networks being highly overparameterized [@dar_farewell_2021]. 
With so many parameters, deep learning models tend to have high variance and thus overfit the training data, a consequence of the bias-variance trade-off common across statistics and machine learning [@mehta_high-bias_2019].
One frequently used method to reduce overfitting is K-fold cross validation [@raschka_model_2020], but this approach cannot be effectively employed when there is one or few time series in the training set.
To remedy the instability problem, we use an ensemble model, wherein each ML forecast is the union of forecasts from 5 individual models that were trained with different initialization seeds.
We found this simple ensemble method to be an effective way to improve generalizability in the limited data regime.

Recently, it has become established that using memory or attention-based neural networks, and an encoder-decoder architecture is crucial for improving forecasting performance on time series data [@kao_exploring_2020; @lyu_lstm_2020; @du_multivariate_2020].
Herein we will provide some background on what these machine learning methods are and their benefits.

### Recurrent Neural Networks

Recurrent neural networks are the predominant memory-based deep learning method.
Recurrent neural networks differ from feed-forward neural networks in that a recurrent neural network provides feedback to itself between time steps [@sherstinsky_fundamentals_2020].
By providing self-feedback, recurrent neural networks are able to retain information from previous time steps and thus learn temporal dependencies.
However, a standard recurrent neural network is unwieldy to train because of the vanishing and exploding gradient problem [@pascanu_difficulty_2013], so there have been specialized neural network architectures designed to avoid these gradient problems.
Long Short-term Memory and Gated Recurrent Units Networks are considered to be the state of the art recurrent neural networks that address exploding and vanishing gradients [@chung_empirical_2014].
These methods avoid gradient problems by regulating the self-feedback between time steps via gates which perform operations on the feedback signal -- see @chung_empirical_2014 for more details.
While GRU's and LSTM's commonly outperform standard RNN's, it is difficult to anticipate whether GRU's or LSTM's will be best suited for any time series problem [@chung_empirical_2014], so we investigate both methods.

### Transformers

The Transformer is a state of the art ML architecture that is able to model long and short term dependencies on sequence to sequence tasks [@vaswani_attention_2017].
Transformers use a mechanism called self-attention which interrelates different positions of the input sequence in order to find an informative representation of the input sequence [@vaswani_attention_2017].
For example, if given a sentence, a transformer could learn the contextual relationship between a subject and a direct object, but a recurrent neural network would process all the words as one phrase.
Because of self-attention, Transformers do not need to process data sequentially and thus can be parallelized, offering significant computational advantages [@vaswani_attention_2017].
The Transformer is likely to be a foundational method for future AI research [@bommasani_opportunities_2021], so we considered it critical to investigate Transformers in this study.

### Encoder-Decoder

Encoder-decoder architectures have been shown empirically to excel on sequence to sequence tasks [@cho_learning_2014].
Encoder-decoders work by processing the input sequence into a fixed-length vector then decoding the fixed-length vector to the predicted output sequence.
It is thought that by encoding the input sequence to a vector, encoder-decoders find informative representations of the input sequence that make the prediction task easier for the decoder [@sutskever_sequence_2014].
Of the models that we present, Block RNNs and Transformers are the only models that use encoder-decoders.

<!--
Basic typology of machine learning methods, how we distinguish "ML" from statistical methods (i.e. bias-variance tradeoff; no criteria to prove that the algorithm produces an unbiased estimate of a given summary statistic)

Stuff about how most machine learning techniques provide only point forecasts
-->

## Forecast skill: strictly proper scores

To compare forecasts, we focus exclusively on metrics of forecast skill which satisfy @gneiting_strictly_2007's property of a strictly proper score.
This ensures the very desirable behavior that which no probabilistic forecast $Q(x,t)$ can have a score as high as that of the true process $P(x,t)$ on average. 
In other words, while it is possible for any of the models considered to _overfit_ the data against which they are trained, i.e. have a higher likelihood than the true process, it is not possible for these models to overfit the data against which they are scored.
It is worth noting that this property applies specifically to probabilistic forecasts and not point forecasts. Not all common metrics metrics often used to compare forecasts are strictly proper -- such as the average root-mean-square error or the average absolute error. <!-- MFL: hmm I am not completely following this point here. CB: added some examples of non-strictly-proper scoring metrics. For instance, {darts} has a whole 'metrics' class, which does not include any strictly proper metrics (because ML researchers don't care about statistical theorems?) -->
Concerns about over-fitting arise in most types of model estimation and are a particularly acute concern to machine learning methods due to the bias-variance trade-off [@mehta_high-bias_2019]. <!-- MFL: I don't think I nailed the overfitting part in ML section but we probably want to reference this here-->
This makes the use of strictly proper scoring especially relevant in assessing machine learning predictions.

Not even all strictly proper scores will agree on the same relative ranking between forecasts.
We will focus on two of the most common such skill metrics, CRPS score and log probability score (negative log likelihood) [e.g. see @gneiting_strictly_2007; @gneiting_probabilistic_2014].
Of the two, the logs probability score puts a much a greater penalty on unexpected observations than CRPS, and may be more suitable when the occurrence of unexpected events incurs a particularly high cost.
Note that while the minus log-likelihood can be negative for sufficiently high probability densities, we use a fixed scalar shift of logs score to ensure the log skill score is strictly positive, which facilitates visualization without impacting relative rankings.

# Results

We examine forecast skill for each of the six forecasting methods (MCMC, ARIMA, block-RNN, GRU, LSTM, and transformer) in each of our three scenarios (Hopf bifurcation, saddle-node bifurcation, stochastic transition).
In addition to these cases, we also consider an "ensemble model", generated by drawing from the distribution of all models except the MCMC model. 
Such ensemble techniques can better reflect uncertainty than relying on any single method [@gneiting_weather_2005].
For simplicity, we consider the unweighted case, where each model is represented equally in the ensemble.
Using model-based simulations allows us to examine performance against a multiple (n=100) replicates of the "true" process, which further helps identify differences that may occur solely due to chance. 
By taking the true model structure as given, MCMC methods can be used to determine a theoretical limit of forecasting skill.
Note that in both bifurcation scenarios, future dynamics will see states never previously observed in the historical data that was used to train each of the methods (i.e. very small population sizes).  
This no-analog aspect of forecasting bifurcation dynamics means that even with many sample points in the training data _and_ perfect knowledge of the true model structure, posterior distributions of parameter values are still influenced by the choice of priors.

```{r}
scores <- arrow::open_dataset("../scores", format="csv") |> collect()
# shift logs score to strictly positive, non-infinite values so we can use a log-scale
shift <- scores |> 
  filter(!is.infinite(logs)) |> 
  mutate(logs = logs - min(logs) +.01) |>
  pivot_longer(c(logs, crps), 
               names_to = "metric", 
               values_to = "score")
```

```{r}
#| fig.cap="Overall distribution of skill scores across models, including an ensemble of methods.
#| Smaller scores are better (indicating smaller errors).
#| Black bars indicate means, dots indicate all individual predictions over time and replicate 'true'
#| simulations of the given scenario."
shift |> 
  ggplot(aes(forecasting_model, score, color=forecasting_model)) + 
  geom_jitter(alpha = 0.06, show.legend = FALSE, shape=".") + 
  stat_summary(fun = mean, shape="|", show.legend = FALSE, col="black") +
  coord_flip() +
  facet_grid(simulation ~ metric, scales = "free") + 
  scale_y_log10() +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

Overall forecasting skill scores for each model across all three scenarios are summarized in figure 2.
Average scores (black lines) hide wide variation in forecast skill.
Generally, ML performance tends to be bracketed between MCMC (essentially the theoretical optimum), and the statistical ARIMA model, though sometimes performing worse than ARIMA or better than MCMC.
Under scenarios with alternative stable states (saddle and stochastic), the distribution of scores is often bimodal for ML models, though not MCMC. 
The ML ensemble model often performs as well as the best ML model on average.
Note that a wide prediction of uncertainty does not mean a wide range in the score skill -- for instance the ensemble model which has the widest array of outcomes often has the a relatively tight distribution of score, especially in logs skill.
This reflects the relative contributions of accuracy and uncertainty as components in the forecasts.
Overall, ML scores are comparable to MCMC skill except for the scenario of the saddle-node bifurcation, where all other models are much worse.

```{r}
forecasts <- read_parquet("data/forecasts.parquet") |> filter(forecasting_model != "ml_ensemble")
observations <- read_parquet("data/observations.parquet")

hopf_obs <- observations |> mutate(variable = as.character(variable)) |> filter(iter %in% 1:15, simulation == "hopf", variable !="X")
saddle_obs <- observations |> filter(iter %in% 1:15, simulation == "saddle")
stochastic_obs <- observations |> filter(iter %in% 1:15, simulation == "stochastic")

```

```{r }
#| fig3
#| fig.cap="Forecasts of the Hopf bifurcation under each model, compared to 15 realizations of the true model. 
#| The bifurcation occurs soon after forecasting period begins,
#| leading to progressively larger and larger oscillations.
#| Prior to the bifurcation, pseudo-cycles are visible in the training data due to stochastic excitations. 
#| Following the bifurcation, stochasticity blurs the oscillatory pattern across replicate simulations.
#| Only last 25 time points of training data are shown."

forecasts  |> 
  filter(iter %in% 1:15, simulation == "hopf", variable !="X") |> 
  ggplot(aes(x=t, y=value, group = interaction(iter, variable), col=type)) + 
  geom_line(alpha=0.1, data = hopf_obs) +
  geom_line(alpha=0.1) +
  geom_line(data = filter(hopf_obs, type=="historical")) +
  facet_grid(forecasting_model ~ variable, scales = "free") +
  theme_bw()  + 
  coord_cartesian(ylim = c(0,15), xlim=c(75, 200)) + 
  ggtitle("Hopf bifurcation")
```

Forecasts of the Hopf bifurcation (Figure 3) are roughly comparable across the phenomenological models (ARIMA and machine learning models).
All models are trained using 100 time points drawn from the period of time prior to the onset of the Hopf bifurcation, which leads to a stable limit cycle that gradually grows in magnitude.
Most models predict a roughly constant mean with a spread roughly equal to that created by the stochastic oscillations around the stable node as seen in the training data prior to the bifurcation.
Notably, the GRU method picks up the oscillatory nature of the dynamics, despite the fact that no true oscillations were yet present in the training data.
However, like the other ML models, it fails to predict the growing amplitude of those oscillations.
Having access to the true model structure, the MCMC model alone predicts the transition into a pattern of oscillations which grows over time, though it tends to overestimate the amplitude of those oscillations initially.
Despite this, overall all methods score comparably in CRPS score (Fig 2), with most ML methods actually out-performing the MCMC score on average (Fig 5), albeit with much greater variation in individual scores.
A clearer picture can be seen by looking at these skill scores over time (Fig 6-7), which show that MCMC is initially performing worse (over-predicting variance) but as oscillations grow further, it starts outperforming the more stationary forecasts of the ML models.

<!-- would be interesting to revisit this with lower stochasticity such that the cyclic behavior is more crisp, and the quasi-cycles in the training data are smaller-->

```{r }
#| fig.cap = "Forecasts of the saddle-node bifurcation under each model, compared to 15 realizations of the true model.
#| Training data proceeds the bifurcation, making accurate prediction without knowledge of the underlying model very difficult."

forecasts |> 
  filter(iter %in% 1:15, simulation == "saddle") |> 
  ggplot(aes(x=t, y=value, group = interaction(iter, variable), col=type)) + 
  geom_line(alpha=0.1, data = saddle_obs) +
  geom_line(alpha=0.1) +
  geom_line(data = filter(saddle_obs, type=="historical")) +
  facet_wrap(~forecasting_model, scales = "free") +
  theme_bw()  + coord_cartesian(ylim = c(0,1)) + 
  ggtitle("Saddle-Node bifurcation")
```

The saddle node bifurcation proves even more difficult for most methods (Figure 4).
Only the MCMC model anticipates the sharp transition to an alternative state, though this behavior is more baked into the method from the start which takes the true model structure as given. 
Even accurate estimation of the MCMC requires slightly informative priors, though still broad enough to reflect a wide range of possible outcomes. 
Two ML methods -- Block RNNs and transformers -- resemble a naive prediction extrapolating the last observed state, failing even to reflect the slow downward trend of the training data.  
LSTM indicates greater uncertainty, while GRU shows very large variability which spans the alternative stable state range. 
With additional tuning, better performance may be possible for these ML models. 
The selected ARIMA model reflects wide uncertainty that is nevertheless not broad enough to span the alternative stable state.
Consequentially, the MCMC estimate easily outperforms the ML models (Fig 2).
The distribution of scores over

```{r}
#| fig.cap = "Forecasts of the stochastic transition under each model, compared to 15 realizations of the true model.
#| In contrast to the other challenges, this case considers the prediction of replicate systems starting from the same initial condition, rather than forecasting the future evolution of the model after the stochastic transition has already occurred."
forecasts |>
  filter(iter %in% 1:15, simulation == "stochastic") |> 
  ggplot(aes(x=t, y=value, group = interaction(iter, variable), col=type)) + 
  geom_line(alpha=0.2, data = filter(stochastic_obs, type=="observed")) +
  geom_line(alpha=0.2) +
  geom_line(data = filter(stochastic_obs, type=="historical")) +
  facet_wrap(~forecasting_model, scales = "free") +
  theme_bw()  + coord_cartesian(ylim = c(0,0.8)) + 
  ggtitle("Stochastic transition")
```
Machine learning methods do markedly better on the stochastic transition scenario than in the two bifurcation scenarios (Fig 5).
This occurs because the training data includes the transition phenomenon of interest.
All ML models accurately capture the dynamics of a sharp transition between alternative stable states -- a dynamic the statistical ARIMA model entirely fails to reflect.
Stochastic transition events should be approximately exponentially distributed, as seen in the wide range of waiting times for transitions to occur in replicates of the true 'observed' process (Fig 5).
Transformer and Block RNN distribution times are much more concentrated, while again GRU and especially the LSTM do a better job reflecting the uncertainty in range of transition times.

```{r}
#| fig6
#| fig.cap="CRPS scores over time for each scenario.
#| Each line represents the scores against a replicate of time series observations from 'true model'.
#| Patterns show forecast skill generally get worse over time -
#| gradually in the case of Hopf bifurcation or suddenly in repsonse to the saddle-node bifurcation.
#| The MCMC shows a sharp improvement in forecast skill over time as it passes the transition window.
#| In the stochastic transition context, this creates two branches, where high values indicate periods of
#| time when the forecast predicts the wrong state and lower branch indicates the correct state."

shift |> 
  group_by(simulation,forecasting_model, metric) |> 
  mutate(t = t - min(t)) |> ungroup() |> 
  filter(metric == "crps", t<100) |> 
  ggplot(aes(t, score, col = forecasting_model, fill=forecasting_model)) + 
  geom_line(aes(group=iter), alpha = 0.05, show.legend=FALSE) +
  facet_grid(simulation ~ forecasting_model, scales="free") + 
  theme_bw()  + ggtitle("CRPS scores over time")
```

```{r}
#| fig.cap="Logs skill score over time. 
#| Forecasts which underestimate uncertainty do substantially worse in logs score than in CRPS score.
#| Comparing this panel to those in Figure 6 highlights scenarios that most often underestimate uncertainty.
#| Generally, MCMC performs better relative to other models under this metric than it does under CRPS,
#| reflecting the bias-variance tradeoff of machine learning."

shift |> 
  group_by(simulation,forecasting_model, metric) |> 
  mutate(t = t - min(t)) |> ungroup() |> 
  filter(metric == "logs", t<100) |> 
  ggplot(aes(t, score, col = forecasting_model, fill=forecasting_model)) + 
  geom_line(aes(group=iter), alpha = 0.05, show.legend=FALSE) +
  facet_grid(simulation ~ forecasting_model, scales="free") + 
  theme_bw() + scale_y_log10() + ggtitle("Logs skill score over time")
```

# Discussion

Ecological systems have long been acknowledged as complex, due not only to the immense span of dimension and scale such processes involve, but also the frequency of emergent and non-linear phenomena such as stochastic resonance, including bifurcations, tipping points, and hysteresis examined here.
Calls for increased forecasting efforts from ecologists frequently reference the role of changing climate and other anthropogenic change, which raise the challenge of prediction in no-analog environments, anticipating ecosystem responses to conditions that have not been previously observed [@clark_ecological_2001; @dietze_iterative_2018].
This motivates the question, "What methods will be most reliable in the face of unobserved conditions?"

In this paper, we carry out an initial exploration on how deep learning methods can perform on predicting transition events.
We compare the ability of several cutting edge machine learning approaches against statistical and process-based models, and show that deep learning methods are able to strike a middle ground between what we consider as acceptable and ideal case forecasting methods, ARIMA and MCMC-based parameter estimation respectively.
When machine learning models are able to observe transition events, as in the stochastic case, they generally outperformed MCMC-based forecasting but under-performed MCMC when there were no transition events in the training sets as in the Hopf and Saddle-node examples.
Yet, there are nuances to these assessments of performance since forecasting skill varies over time as we show in @figureX.
For instance, while most of the ML methods outperform MCMC in the stochastic transition example when forecasting skill is averaged over all time points, MCMC produces more replicates with relatively small errors than the ML models.

There are a number of questions that we have left unanswered in our study that we hope will be addressed in future work.
In this paper, we have explored a small number of machine learning and statistical models that can be used for forecasting, so general conclusions can not be drawn on whether statistical or machine learning-based approaches are better suited for tipping point forecasting problems.
Neither can we claim that the ML methods employed will translate well to all transition event forecasting  problems in reality, since working with real data will introduce additional difficulties.
For instance, to train a neural network on more than one real time series, we would likely have to employ a specialized algorithm that would allow for generalization over varying time and state scales.

Our analysis has focused on the task of making a single forecast prior to the occurrence of a critical transition.
Forecasting is ideally a more iterative process of data assimilation, where forecasts are updated with respect to additional observations, rather than projecting 100s of time steps into the future [@Dietz2018]. 
Updating a forecast after a critical transition has already occurred may be of little use in the context of hysteresis, such as under the saddle node or stochastic transition -- recognizing the alternative stable state only after the system is stuck in that basin will often be considered 'too late'.
Assimilation may be more applicable to the Hopf bifurcation, where additional observations of slowly growing oscillations may lead to more accurate forecasts. 
Such models may even accurately predict the homoclinic bifurcation that occurs when the limit cycle grows too large, eventually hitting a saddle point of zero population size for the host species.
We leave these cases to future exploration.

Ecological forecasting is invariably difficult, even in the idealized cases of ample measurement data and clearly identified structural models.
This paper is not intended to give a comprehensive answer to whether machine learning models are the the best methods to use on any tipping point forecasting problem as this will take numbers of studies to resolve. 
Instead, this paper aims to serve as a first exploration on whether deep learning methods should be considered as viable tools for the extremely challenging category of no-analog bifurcation forecasting problems.

- overall forecasting bifurcations is difficult, even in the idealized case of ample measurement data and knowing the structural model exactly.

- ML techniques are mostly bracketed by ARIMA and MCMC as expected (better than arima, worse than estimates using true model), though details vary (can do better than MCMC etc).

- MCMC is most relevant to the saddle-node, which is an especially hard problem for the model-free methods.  block rnn does surprisingly well here all things considered.

- the stochastic context is in some sense the easiest for  ML/statistical methods, as it is the only case where training data includes the phenomenon of the transition.

- MCMC does not always out-perform ML techniques, particularly in CRPS terms. Log likelihood-based skill score

- Forecast skill assessment can be challenging. it is essential to pay attention to skill over time. Forecast error is not a smooth function of horizon, nor is it always increasing with longer horizons.

- Examining means and summary statistics of variation (confidence intervals) can be limiting too, obscuring aspects such as bimodally distributed scores.

- Forecasts with 'assimilation' i.e. ability to re-estimate after seeing some of the data, rather than looking at a single forecast with a long horizon.

\pagebreak

# References